### 2023-08-14 paper/main.tex
final_paper

@@

According to the sampling theorem, spatial aliasing can be avoided by analytically spatial-bandlimiting the driving functions to the Nyquist wavenumber.
This can be performed by either spatially filtering the driving functions with an appropriate spatial low-pass filter prior to numerical evaluation, or performing bandlimitation directly in the wavenumber (modal) domain analytically.
The latter approach is implemented by \cite{hahn2022cylindrical}, and will serve as a reference solution in the following investigation.
However, the modal bandlimitation solution is available only for plane wave and point source virtual source models, at the cost of increased computational complexity.

@@

The shifted spatially filtered driving function reads as
\begin{multline}
    d'_x(s,t+ \tau(s) ) = \int h_x(s-s_0) \, d(s_0,t+ \tau(s) ) \, \td s_0 = \\
    \int h_x(s-s_0) \, A^D(s_0)\, \delta(s_0,t+ \left(\tau(s)-\tau(s_0)\right)  ) \, \td s_0.
    \label{eq:spat_conv2}
\end{multline}
In order to evaluate the spatial convolution the generalized sifting property of the Dirac-delta may be applied \cite{Firtha2019phd, Jackson1999}, which states that
\begin{equation}
    \int f(s_0) \delta(g(s_0)) \td s_0 = \sum_i \frac{f(s_i)}{| \frac{\partial}{\partial s_0} g(s_0) |_{s = s_i}}, \hspace{3mm} g(s_i) = 0.
\end{equation}
In the present problem the zeros of the Dirac's argument are found where
\begin{equation}
    \tau(s_i) = t + \tau(s)
    \label{eq:zero}
\end{equation}
is satisfied.
In the following it is assumed that a single zero $s_i$ exists, satisfying \eqref{eq:zero}.
This assumption means that the wavefront arrives at each SSD element at a unique time instant, being strictly true for a virtual plane wave and a linear SSD.
The derivative of the phase function is given by
\begin{equation}
    \frac{\partial}{\partial s_0} g(s_0) = - \frac{\partial}{\partial s_0} \tau(s_0) = -\frac{\hat{k}_t^P(s_0)}{c}.
\end{equation}
Here $\hat{k}_t^P(s_0)$ is the tangential component of the local propagation vector, defined as
\begin{equation}
    \hat{k}_t^P(s_0) = \mathbf{t}^{\mathrm{T}}(s_0) \hat{\vk}^P(s_0),
\end{equation}
with $\mathbf{t}^{\mathrm{T}}(s_0)$ denoting the unity long tangential vector of the SSD element at position $s_0$.

@@

Since the cutoff frequency of the antialiasing filter (Eq. \eqref{eq:cutoff_fr}) is inversely proportional to the tangential component of the virtual field on the SSD, in a circular array a single SSD element will have nearly fullband driving signal: that SSD element at which the wavefront arrives perpendicularly (i.e. where the tangential component of the local propagation vector is zero).
This fullband SSD element is marked with white dot in Figure \ref{Fig:symm_antialiasing} (a).
From general WFS theory it is known that the fullband SSD element will dominate the synthesized wavefront (i.e. will serve as a stationary SSD position) for spatial positions along a straight line crossing the actual SSD element into the direction of the local propagation vector.
Therefore, the present antialiasing strategy ensures aliasing-free synthesis along a straight line for which the fullband SSD element is the stationary secondary source.
The direction of the line of aliasing-free synthesis is marked with a dashed arrow in the figure.
This is in contrast with previous LWFS approaches, in which the synthesis is optimized in the proximity of a pre-defined reference position over a circular area.

>>>>

\usepackage{amsmath}
\usepackage{bm}
\usepackage{tkz-euclide}
\usetikzlibrary{shapes,snakes}
% we should use matplotlib colors:
\definecolor{C0}{HTML}{1f77b4}
\definecolor{C1}{HTML}{ff7f0e}
\definecolor{C7}{HTML}{7f7f7f}

\makeatletter
\newcommand{\tikzAngleOfLine}{\tikz@AngleOfLine}
\def\tikz@AngleOfLine(#1)(#2)#3{%
  \pgfmathanglebetweenpoints{%
    \pgfpointanchor{#1}{center}}{%
    \pgfpointanchor{#2}{center}}
  \pgfmathsetmacro{#3}{\pgfmathresult}%
}
\makeatother
\tikzset{mark coordinate/.style={inner sep=0pt,
                                   outer sep=0pt,
                                   minimum size=3pt,
                                   fill=#1,
                                   circle}
                                   }

@@

According to the sampling theorem, spatial aliasing can be avoided by analytically spatial-bandlimiting the driving functions to the Nyquist wavenumber.
This can be performed by either spatially filtering the driving functions with an appropriate spatial low-pass filter prior to numerical evaluation, or performing bandlimitation directly in the wavenumber (modal) domain analytically.
The latter approach is implemented by \cite{hahn2022cylindrical}, and will serve as a reference solution in the following investigation.
However, the modal bandlimitation solution is available so far only for plane wave and point source virtual source models---where the driving functions can be derived analytically---, at the cost of increased computational complexity.

@@

The shifted spatially filtered driving function reads as
\begin{multline}
    d'_x(s,t+ \tau(s) ) = \int h_x(s-s_0) \, d(s_0,t+ \tau(s) ) \, \td s_0 = \\
    \int h_x(s-s_0) \, A^D(s_0)\, \delta(s_0,t+ \left(\tau(s)-\tau(s_0)\right)  ) \, \td s_0.
    \label{eq:spat_conv2}
\end{multline}
In order to evaluate the spatial convolution the generalized sifting property of the Dirac-delta may be applied \cite{Firtha2019phd, Jackson1999}, which states that
\begin{equation}
    \int f(s_0) \delta(g(s_0)) \td s_0 = \sum_i \frac{f(s_i)}{| \frac{\partial}{\partial s_0} g(s_0) |_{s = s_i}}, \hspace{3mm} g(s_i) = 0.
\end{equation}
In the present problem the zeros of the Dirac's argument are found where
\begin{equation}
    \tau(s_i) = t + \tau(s)
    \label{eq:zero}
\end{equation}
is satisfied.
In the following it is assumed that a single zero $s_i$ exists, satisfying \eqref{eq:zero}.
This assumption means that the wavefront arrives at each SSD element at a unique time instant, being strictly true for a virtual plane wave and a linear SSD.
The derivative of the phase function is given by
\begin{equation}
    \frac{\partial}{\partial s_0} g(s_0) = - \frac{\partial}{\partial s_0} \tau(s_0) = -\frac{\hat{k}_t^P(s_0)}{c}.
\end{equation}
Here $\hat{k}_t^P(s_0)$ is the tangential component of the local propagation vector, defined as
\begin{equation}
    \hat{k}_t^P(s_0) = \mathbf{t}^{\mathrm{T}}(s_0) \hat{\vk}^P(s_0),
\end{equation}
with $\mathbf{t}^{\mathrm{T}}(s_0)$ denoting the unity long tangential vector of the SSD element at position $s_0$, as illustrated in Figure \ref{fig:tangential_local_prop_vec}.

\begin{figure}
    \begin{center}
        \begin{tikzpicture}[scale=1.2 ]
    \begin{scope}

    % these coordinates / parameters can be freely chosen:
    \coordinate (x0) at (-1.5,0);  % determines the radius of SSD
    \coordinate (xPS) at (-3.5,2);  % make sure that point source is outside of SSD
    % for wave pattern 'snake=expanding waves' stuff
    \def \segmentangle {30};  % opening angle of wave pattern in deg
    \def \segmentlength {5};  % somehow determines the wavelength
    \def \wavepatternlength {3.5};  % determines how far the wave propagates into the SSD

    % ---> change with care:
    \coordinate (origin) at (0,0);
    % SSD circle
    \node[name path=FullSSDCircle, draw, thin, color=black, circle through=(x0)] (FullSSDCircleNode) at (origin) {};
    % wave pattern to x0
    \draw[snake=expanding waves, segment angle=\segmentangle, segment length=\segmentlength, color=C7!33] (xPS)  -- (x0);
    % wave pattern additionally into SSD, length is defined with \wavepatternlength
    \tikzAngleOfLine(x0)(xPS){\angle};
    \draw[snake=expanding waves, segment angle=\segmentangle, segment length=\segmentlength, color=C7!33] (xPS)  -- ++(\angle:-\wavepatternlength);
    % line from point source to specific SSD point
    \draw[-, thick, C7] (xPS) -- (x0) node[left, black]{$\bm{x}_0$};
    % normal / tangent vectors of SSD at x0
    \tikzAngleOfLine(origin)(x0){\angle};
    \draw[->, >=stealth', thick, C0] (x0) -- ++(\angle:-1) node[right]{$\bm{n}$};
    \draw[->, >=stealth', thick, C1] (x0) -- ++(\angle-90:-1) node[below]{$\bm{t}$};
    % normal / tangent vectors of sound field at x0
    \tikzAngleOfLine(x0)(xPS){\angle};
    \draw[->, >=stealth', thick, C1] (x0) -- ++(\angle:-1) node[right]{$\hat{\bm{k}}$};
    \draw[->, >=stealth', thick, C0] (x0) -- ++(\angle+90:-1) node[right]{$\hat{\bm{k}}_t$};
    % black dots indicating positions of point source and specific SSD point
    \draw[thick] (xPS) coordinate [mark coordinate=black, label=left:{$\bm{x}_\text{PS}$}];
    \draw[thick] (x0) coordinate [mark coordinate=black];
    % <---

    % !!!
    % these projection lines are hard coded for (x0) at (-1.5,0); (xPS) at (-3.5,2);
    \draw[dashed, C0] (-0.79,0.7) -- (-0.79,0) node[below] {$\hat{k}_t$};
    \draw[dashed, C1] (-0.8,-0.7) -- (-1.5,-0.7) node[left] {$\hat{k}_t$};

    \end{scope}
    \end{tikzpicture}
\end{center}
\caption{Illustration for the tangential component of the local propagation vector in the exemplary case of the field of a virtual point source, located at $\vx_{\mathrm{PS}}$ and applying a circular SSD.
On the figure $\bm{t}$ and $\bm{n}$ denote the tangential and normal vector of the SSD respectively, $\hat{\bm{k}}_t$ denotes the tangential vector of the sound sound field and $\hat{k}_t$ denote the tangential component of the local propagation vector ($\hat{\bm{k}}$) at the SSD position $\vxo$.}
\label{fig:tangential_local_prop_vec}
\end{figure}

@@

Since the cutoff frequency of the antialiasing filter (Eq. \eqref{eq:cutoff_fr}) is inversely proportional to the tangential component of the virtual field on the SSD, in a circular array a single SSD element will have nearly fullband driving signal: that SSD element at which the wavefront arrives perpendicularly (i.e. where the tangential component of the local propagation vector is zero).
This fullband SSD element is marked with white dot in Figure \ref{Fig:symm_antialiasing} (a).
From general WFS theory it is known that the fullband SSD element will dominate the synthesized wavefront (i.e. will serve as a stationary SSD position) for spatial positions along a straight line crossing the actual SSD element into the direction of the local propagation vector \cite{Firtha2019phd}.
Therefore, the present antialiasing strategy ensures aliasing-free synthesis along a straight line for which the fullband SSD element is the stationary secondary source.
The direction of the line of aliasing-free synthesis is marked with a dashed arrow in the figure.
This is in contrast with previous LWFS approaches, in which the synthesis is optimized in the proximity of a pre-defined reference position over a circular area.
